First, I want to thank my advisors, Greg Turk and Karen Liu. It is a great pleasure to work with you two in the last six years. You serve as great role models not only in research but also in life. I feel that I am deeply indebted to both of you. You have built up a huge portion of knowledge that I possessed today. You encouraged me to pursue my interests and to follow my passion. You taught me how to think creatively and helped me develop rigorous procedures to approach research problems. More importantly, you helped me build my confidence, especially in my first years in US when I was facing language and culture barriers. I feel extremely fortunate to have you as my advisors. Without your guidance, I would be nowhere as successful as I am today. I believe that what I have learned from you in the last six years will carry on to influence me and benefit me throughout my whole life.

I want to thank the members of my thesis committee: Jarek Rossignac, Frank Dellaert and James O'Brien as well as my qualifier committee members: Irfan Essa and Blair Macintyre. Each of you has helped me become a stronger researcher in your own way. You have broaden my knowledge, shaped my rigorous thinking, taught me effective ways of communication and helped me develop social skills.

I also want to thank my friends and labmates: Huamin Wang, Chris Wojtan, Sumit Jain, Yuting Ye, Karthik Raveendran, Mark Luffel, Tina Zhuo, Yunfei Bai, Sehoon Ha, Kristin Siu, Albert Li, Yangfeng Ji, Alexander Zook, Alex Clegg, Edward Liu, and many more. Thank you for your helpful discussions with me about my research, my presentation and my life.

Finally, I want to thank my family. I want to thank my wife, Yuting, for your unlimited support every day throughout my PhD journey. I also want to thank you for your understanding, especially for your help on modeling, rendering and video editing before each SIGGRAPH deadline. I want to thank my parents, Yangsun and Xiuchun. It is you that shaped me into who I am today. I am extremely grateful to your guidance and supports.
